---
title: "Difference Scores: Paired Samples T-test"
author: "Biostatistics Working Group"
---

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered

# This code configures knitr code chunk options

knitr::opts_chunk$set(
    echo = T, message = F, warning = F, error = F, output = T,
    comment = NA, cache = T,
    R.options = list(width = 220, digits = 3),
    fig.align = "center",
    out.width = "75%", fig.asp = .75
)
```

# Height Example

# Load necessary libraries
```{r}
library(mosaic)
library(ggplot2)
library(WRS2)
library(lm.beta)
library(gamm4)
library(dplyr)
library(plyr)
library(rstatix)
```


# Set seed for reproducibility
```{r}
set.seed(123)
```

# Define means and standard errors (retrieved from https://www.cdc.gov/nchs/data/series/sr_03/sr03-046-508.pdf)

```{r}
mean_male <- 175.3 # for all groups 20+
mean_female <- 161.3 # for all groups 20+
se_male <- 0.19 #n = 5,092
se_female <- 0.19 #n = 5,510
sd_male <- 13.6
sd_female <- 14.1
n_male <- 5092  # sample size
n_female <- 5510 # sample size
samp_n <- 30
```


# Generate sample height data for males in cm
```{r}
heights_male <- rnorm(n = samp_n, mean = mean_male, sd = sd_male) #use estimates from larger sample, pull n = samp_n
```

# Generate sample height data for females in cm
```{r}
heights_female <- rnorm(n = samp_n, mean = mean_female, sd = sd_female) #use estimates from larger sample, pull n = samp_n
```

# Combine data into a data frame
```{r}
height_data <- data.frame(
  Sex = factor(c(rep(0, samp_n), rep(1, samp_n))),  # Convert to numeric: 1 for Male, 0 for Female
  Height = c(heights_female, heights_male)
)

# Perform the t-test
t_test_result <- t.test(Height ~ Sex, data = height_data, var.equal = TRUE)
# Compute Cohen's d
cohen_d<-cohens_d(Height ~ Sex, data = height_data, var.equal = TRUE)

# Generate sample height data for males for larger sample in cm
heights_male_larger <- rnorm(n = n_male, mean = mean_male, sd = sd_male) #use estimates from larger sample

# Generate sample height data for females in cm
heights_female_larger <- rnorm(n = n_female, mean = mean_female, sd = sd_female) #use estimates from larger sample


# Combine data into a data frame
height_data_larger <- data.frame(
  Sex = factor(c(rep(0, n_female), rep(1, n_male))),  # Convert to numeric: 1 for Male, 0 for Female
  Height = c(heights_female_larger, heights_male_larger)
)

# Perform the t-test
t_test_result_larger <- t.test(Height ~ Sex, data = height_data_larger, var.equal = TRUE)
# Compute Cohen's d
cohen_d_larger<-cohens_d(Height ~ Sex, data = height_data_larger, var.equal = TRUE)


# Extract the means and standard deviations from the t-test result
means_male <- t_test_result$estimate[1]
means_female <- t_test_result$estimate[2]

# Subset data for males and females
male_heights <- height_data$Height[height_data$Sex == 1]
female_heights <- height_data$Height[height_data$Sex == 0]


# Reverse the order of levels in the Species variable
height_data$Sex <- factor(height_data$Sex, levels = levels(height_data$Sex))

# Define Ravenclaw inspired colors
ravenclaw_colors <- c("#222f5b", "#b0b7bc")  # Blue and Silver
```
# Plot the barplot with error bars
```{r}
p_bar <- ggplot(height_data, aes(x = Sex, y = Height, fill = Sex)) +
  geom_bar(stat = "summary", fun = "mean", color = "black", width = 0.4) +  # Adjust width here
  geom_errorbar(stat = "summary", fun.data = "mean_cl_boot", width = 0.1, color = "black") +
  geom_text(x = 0.5, y = max(height_data$Height) + 3, 
  label = paste("t(", t_test_result$parameter, ") = ", abs(round(t_test_result$statistic, 2)), ", p = ", round(t_test_result$p.value, 3), sep = ""), 
  size = 6, color = "black", vjust = 0, hjust = -.55) +
  labs(x = "Sex", y = "Mean Height (cm)", title = "Mean Height by Sex") +
  scale_fill_manual(values = ravenclaw_colors, labels = c("Female", "Male")) +  # Use Ravenclaw colors
  scale_x_discrete(labels = c("Female", "Male")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,max(height_data$Height) + 15)) +
  theme_bw() +  # Set background theme to black and white
  theme(text = element_text(size = 18),  # Increase text size
        panel.border = element_blank(),  # Remove panel border
        panel.grid.minor = element_blank(),  # Remove major grid lines
        axis.line.x = element_line(), axis.line.y = element_line())
```


# Save the barplot as a TIFF file with dimensions 9 x 6 inches
```{r}
ggsave(filename = "barplot_height_by_sex.tiff", plot = p_bar, width = 9, height = 6, dpi = 600)
```

# Perform linear regression
```{r}
lm_model <- lm(Height ~ Sex, data = height_data)
print("\nResults of linear regression:")
summary(lm_model)

# Compute correlation coefficients and p-values
corr_result <- cor.test(height_data$Height, as.numeric(height_data$Sex))
print("Results of correlation test:")
print(corr_result)
```

# Plot the scatterplot with linear regression line
```{r}
p_scatter <- ggplot(height_data, aes(x = as.numeric(Sex), y = Height, color = Sex)) +
geom_point(alpha = 0.6, position = position_jitter(width = 0.02, height = 0), size = 3) +  # Larger dots
geom_smooth(method = "lm", se = TRUE, color = "black") +
geom_text(x = 0.5, y = max(height_data$Height) + 3,
label = paste("β = ", round(lm.beta(lm_model)$standardized.coefficients[2], 2), ", t(", lm_model$df.residual, ") = ", round(summary(lm_model)$coefficients[2, 3], 2),
", p = ", round(summary(lm_model)$coefficients[2, 4], 3), sep = ""),
vjust = -0.5, hjust = -0.25, size = 6, color = "black") +
labs(x = "Sex", y = "Height (cm)", title = "Height by Sex") +
scale_color_manual(values = ravenclaw_colors, labels = c("Female", "Male")) +  # Use Ravenclaw colors
scale_x_discrete(labels = c("Female", "Male")) +
coord_cartesian(ylim = c(150, max(height_data$Height) + 10)) +  # Adjust y-axis limits
theme_bw() +  # Set background theme to black and white
  theme(text = element_text(size = 18),  # Increase text size
        panel.border = element_blank(),  # Remove panel border
        panel.grid.minor = element_blank(),  # Remove major grid lines
        axis.line.x = element_line(), axis.line.y = element_line())
```


# Save the scatterplot as a TIFF file
```{r}
ggsave(filename = "scatterplot_height_by_sex.tiff", plot = p_scatter, width = 9, height = 6, dpi = 600)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
